---
title: "bellabeat casestudy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# summary 
Bellatbeat, a high-tech company that manufactres health-focused smart products for women. Bellabeat is a successful small company, but they have the potential to become a larger player in the global smart device market Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company.

# 1.Ask 
  
  - what are some trends in smart device usage?  
  
  - How could these trends apply to Bellabeat customers?
  
  - How could these trends help influence Bellabeat marketing strategy?
  
  **Stakeholders** 
  
  - Urška Sršen - Bellabeat cofounder and Chief Creative Officer

  - Sando Mur - Bellabeat cofounder and key member of Bellabeat executive team 
  
  - Bellabeat Marketing Analytics team
  
# 2. Prepare 
  
  **about data**
   
  - The data set is from Fitbit Fitness Tracker Data. The data contains personal fitness tracker from thirty fitbit users, including output for physical activity, heart rate, and sleep monitoring. It inlcudes information about daily activity, steps, and herat rate that  can be used to explore user's habits.
  
# 3. Process 

 - considering accessibility, amount of data and various visualization, I will use R to analyze the data 
 
 - Installing Packages and Opening Libraries 
```{r, echo=TRUE, results = 'hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(skimr)
library(janitor)
library(dplyr)
library(lubridate)
```

importing Dataset
  daily activity, daily calories, hourly steps, daily intensities, daily sleep and weight info
```{r, echo=TRUE, results = 'hide', warning=FALSE, message=FALSE}
activity <- read_csv("Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
Calories <- read.csv("Fitabase Data 4.12.16-5.12.16/dailyCalories_merged.csv")
hourly_steps <- read.csv("Fitabase Data 4.12.16-5.12.16/hourlySteps_merged.csv")
intensities <- read.csv("Fitabase Data 4.12.16-5.12.16/dailyintensities_merged.csv")
sleep <- read.csv("Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
weight <- read.csv("Fitabase Data 4.12.16-5.12.16/weightLogInfo_merged.csv")
```

checking sample size by participants ID
```{r, echo=TRUE, results = 'hide', warning=FALSE, message=FALSE}
n_distinct(activity$Id)
n_distinct(Calories$Id)
n_distinct(sleep$Id)
n_distinct(weight$Id)
n_distinct(hourly_steps$Id)
```
33

33

24

8

33

because the sample size of Weight data is too small for minimum of 30, I'll exclude the dataset for later analyze. For the sleep data, i will still keep the dataset but low validity compare to other datasets.


checking data columns 
 ```{r, eval=FALSE}
colnames(activity)
colnames(intensities)
colnames(sleep)
colnames(hourly_steps)
```


Changing Date format and dropping redundant columns
```{r, echo=TRUE, results = 'hide', warning=FALSE, message=FALSE}
hourly_steps <- hourly_steps %>%  separate(ActivityHour, c("Date", "Hour"), sep = ' ')
activity <- subset(activity, select= -c(SedentaryActiveDistance, LoggedActivitiesDistance))
```

let's have statistic summary tables
```{r}
#summarizing data
#Activity
activity %>%  select(TotalSteps, TotalDistance, VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes, SedentaryMinutes, Calories, ActivityDate) %>% drop_na() %>% summary()
#Calories
Calories %>%  select(Calories) %>% drop_na() %>% summary()
#Sleep
sleep %>% select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>% drop_na() %>% summary()
```

Create some new data frames before get in to Analyze

```{r}
#on average, during which hour of the day were the most steps taken? 
hourly_steps %>% group_by(Hour) %>% summarize(mean_steps = mean(StepTotal)) %>% select(Hour, mean_steps) %>% arrange(desc(mean_steps)) %>%  head(5)
mean_steps <- hourly_steps %>% group_by(Hour) %>% summarize(mean_steps = mean(StepTotal)) %>% select(Hour, mean_steps) %>% arrange(desc(mean_steps)) %>% as.data.frame
head(mean_steps)
#average amount of sleep for each participant 
mean_sleep <- sleep %>% group_by(Id) %>%  summarize(mean_sleep = mean(TotalMinutesAsleep)) %>% select(Id, mean_sleep)
head(mean_sleep)
```

# 4. Analyze & Visualize

**Findings from the summary**

- Daily average sedentary time is 991 minutes (16.5hour). 

- Daily average steps taken: 7638.

- In average, participants spends 458 minutes(7.6hours) time in bed and 419minute(7hour) for actual sleep.

- Most of the participants are lightly active 


What time in a day participants are most active?

```{r}
#average steps per hour 
ggplot(mean_steps, aes(x= Hour, y= mean_steps)) + 
  geom_col(aes(reorder(Hour, +mean_steps), mean_steps))+
  theme(axis.text.x = element_text(angle=90))+  
  labs(title = " Average steps taken per hour of day ")
```

The graph shows that on average, participants are active between 6:00~9:00pm

```{r}
#aeverage munutes of sleep 
ggplot(mean_sleep, aes(x = Id, y = mean_sleep)) + 
geom_col(aes(reorder(Id, + mean_sleep), y = mean_sleep)) + 
labs(title = "Average Minutes of Sleep") + 
theme(axis.text.x =  element_text(angle = 90)) +
geom_hline(yintercept = mean(mean_sleep$mean_sleep), color = "red")
```

box plot for average steps per day


```{r}
ggplot(data=activity, aes(x = TotalSteps)) + geom_boxplot()

activity %>% select(TotalSteps) %>% summary()
```

Daily average steps is 7638 and in range of 4000 to 11000 steps per day

Let's find out correlation between steps and calories 

```{r}
ggplot(data=activity, aes(x=TotalSteps, y=Calories)) + geom_point()+ geom_smooth(color = "red") + labs(title = "total steps vs calories")
correlation <- activity %>% select(TotalSteps, Calories) %>% drop_na() %>% summarize(correlation = cor(TotalSteps, Calories)) 
print(correlation)
```
The result shows that correlation value for Total steps and Calories is 0.592 which is strong correlation

# 5. Conclusion 

1. we found that majority of the participants spent their day in sedentary and mostly active after 7:00pm. Bellabeat should focus their main customer to a women, full time working, especially  those who work sedentary. 

2. Bellabeat could provide periodic reminders for those who sits for prolonged periods to stand up, stretch, and engage in light exercises. It could also offer guidance on effective workouts. 

3. Current medical guidelines suggest that most adults should aim for about 10,000 steps perday. since participants walks about 7500steps daily on average, bellabeat should encourage users to be more active than usual. 


